Deep Learning

### 2015-01-13 deepreview.tex
fixed the missing references

Deep learning is currently making major advances in solving problems
that have resisted the best attempts of the Artificial Intelligence
community for many years. Deep learning is very good at discovering
intricate structure in large, high-dimensional datasets and is
therefore applicable to many domains of science and business. For
example, it has beaten previous machine learning techniques at
predicting the activity of potential drug
molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
particles~\citep{YANNparticles}, classifying
galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations
in non-coding DNA on gene expression and
disease~\citep{Keung-et-al-2014}. % Frey

@@

\iffalse
\begin{figure}
\begin{center}
  \vspace{0.2cm}
  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
  nice convnet figure goes here
\end{center}
\vspace{-0.6cm}
\caption{A typical ConvNet architecture}
\vspace{-0.8cm}
\label{fig:convnet}
\end{figure}
\fi

@@

The architecture of a typical ConvNet is shown in
figure~\ref{fig-convnet}. The first few stages are composed of two
types of layers: convolutional layers and pooling layers. Units in a
convolutional layer are organized in {\em faeture maps}, within which
each unit is connected to local patches in the feature maps of the
previous layer through a set of weights called a {\em filter
  bank}. The result of this local weight sum is then passed through a
non-linearity.  All units in a feature map share the same filter
bank. Multiple feature maps in a layer use different filter banks. The
reason for this architecture is twofold. First, In array data such as
images, local groups of values are often highly correlated forming
local motifs that are easily detected. Second, the local statistics of
images and other signals are invariant to location. In other word, if
a motif can appear in one part of the image, it could appear anywhere,
hence the idea of units at different locations sharing the same
weights.  Mathematically, the filtering operation performed by a
feature map is a {\em discrete convolution}, hence the name
ConvNet. While the role of the convolutional layer is to detect local
conjunctions of features in the previous layer, the role of the
pooling layer is to merge semantically similar features into
one. Since the relative positions of the features forming a motif can
vary somewhat, reliably detecting the motif can be done by
coarse-graining the position of each feature. A typical pooling unit
computes the max of a local patch of units in one feature map (or in a
few). Neighboring pooling units take input from patches that are
shifted by more than one row or column, thereby reducing the dimension
of the representation and creating an invariance to small shifts.  Two
or three stages of convolutionas, non-linearity, and pooling are
stacked, followed by more convolutional and fully-connected layers.
Back-propagating gradients through a ConvNet is as simple
as through a regular deep network, allowing all the weights in all the
filter banks to be trained.

This architecture exploits the property that many natural signals are
{\em compositional hierarchies}. In images, local combinations of
edges form motifs, motifs assemble into parts, and parts form objects.
Similar hierarchies exists in speech and text from sounds to phones,
phonemes, syllables, words, and sentences. The elements that form
elements in the next level can vary in relative position and
appearance. The convolutional and pooling layers in ConvNets are
directly inspired by the classic notions of simple cells and complex
cells in visual neuroscience~\cite{Hubel62}, and the overall
architecture is reminiscent of visual cortex ventral pathway hierarchy
LGN-V1-V2-V4-IT~\cite{VanEssen}. ConvNets have their roots in
Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
somewhat similar but did not have an end-to-end supervised learning
algorithm such as backpropagation+SGD.

@@

Since the early 2000s, there have been many tasks in which ConvNets
have been applied with success to the detection, segmentation, and
recognition of objects and regions in images. These were invariably
tasks in which labeled data was relatively abundant, such as traffic
sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
segmentation of biological images~\cite{ning-05} particularly for
connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,taylor-xx,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.

@@

Other applications of ConvNets that are gaining importance concern
natural language understanding~\citep{collobert:2011b}, and speech
recognition~\cite{sainath-xx}.

>>>>

Deep learning is currently making major advances in solving problems
that have resisted the best attempts of the Artificial Intelligence
community for many years. Deep learning is very good at discovering
intricate structure in large, high-dimensional datasets and is
therefore applicable to many domains of science and business. For
example, it has beaten previous machine learning techniques at
predicting the activity of potential drug
molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
particles~\citep{Melis-Higgs-boson-competition-2014}, % TODO ~\citep{YANNparticles}, 
classifying galaxies, % %TODO ~\citep{YANNgalaxies}, 
and predicting the effects of mutations
in non-coding DNA on gene expression and
disease~\citep{Keung-et-al-2014}. % Frey

@@

%\begin{figure}
%\begin{center}
%  \vspace{0.2cm}
%  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
%  nice convnet figure goes here
%\end{center}
%\vspace{-0.6cm}
%\caption{A typical ConvNet architecture}
%\vspace{-0.8cm}
%\label{fig:convnet}
%\end{figure}

@@

The architecture of a typical ConvNet is 
%shown in
%figure~\ref{fig-convnet}. 
structured as a series of stages. The first few stages are composed of two
types of layers: convolutional layers and pooling layers. Units in a
convolutional layer are organized in {\em faeture maps}, within which
each unit is connected to local patches in the feature maps of the
previous layer through a set of weights called a {\em filter
  bank}. The result of this local weight sum is then passed through a
non-linearity.  All units in a feature map share the same filter
bank. Multiple feature maps in a layer use different filter banks. The
reason for this architecture is twofold. First, In array data such as
images, local groups of values are often highly correlated forming
local motifs that are easily detected. Second, the local statistics of
images and other signals are invariant to location. In other word, if
a motif can appear in one part of the image, it could appear anywhere,
hence the idea of units at different locations sharing the same
weights.  Mathematically, the filtering operation performed by a
feature map is a {\em discrete convolution}, hence the name
ConvNet. While the role of the convolutional layer is to detect local
conjunctions of features in the previous layer, the role of the
pooling layer is to merge semantically similar features into
one. Since the relative positions of the features forming a motif can
vary somewhat, reliably detecting the motif can be done by
coarse-graining the position of each feature. A typical pooling unit
computes the max of a local patch of units in one feature map (or in a
few). Neighboring pooling units take input from patches that are
shifted by more than one row or column, thereby reducing the dimension
of the representation and creating an invariance to small shifts.  Two
or three stages of convolutionas, non-linearity, and pooling are
stacked, followed by more convolutional and fully-connected layers.
Back-propagating gradients through a ConvNet is as simple
as through a regular deep network, allowing all the weights in all the
filter banks to be trained.

This architecture exploits the property that many natural signals are
{\em compositional hierarchies}. In images, local combinations of
edges form motifs, motifs assemble into parts, and parts form objects.
Similar hierarchies exists in speech and text from sounds to phones,
phonemes, syllables, words, and sentences. The elements that form
elements in the next level can vary in relative position and
appearance. The convolutional and pooling layers in ConvNets are
directly inspired by the classic notions of simple cells and complex
cells in visual neuroscience~\cite{Hubel62}, and the overall
architecture is reminiscent of visual cortex ventral pathway hierarchy
LGN-V1-V2-V4-IT~\cite{Felleman+VanEssen-1991}. ConvNets have their roots in
Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
somewhat similar but did not have an end-to-end supervised learning
algorithm such as backpropagation+SGD.

@@

Since the early 2000s, there have been many tasks in which ConvNets
have been applied with success to the detection, segmentation, and
recognition of objects and regions in images. These were invariably
tasks in which labeled data was relatively abundant, such as traffic
sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
segmentation of biological images~\cite{ning-05} particularly for
connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
nasse-09,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.

@@

Other applications of ConvNets that are gaining importance concern
natural language understanding~\citep{collobert:2011b}, and speech
recognition~\cite{Sainath-et-al-ICASSP2013}.
